A quantitative framework for investigating the reliability of network construction: supplemental information

### 2018-06-07 manuscript/supplemental.tex
mostly re-ordered

% Definite first
\section*{Appendix S1: Calculating uninformative and informative priors}

@@

        The mean of these interaction probabilities is 0.176, approximately two-thirds the $\lambda_{MLE}$ obtained from the observed data. We can use the distribution of these interaction probabilities as our prior distribution and estimate the uncertainty surrounding our $\lambda_{MLE}$. With some simple R code (function ``calculate\_parameters", \emph{Appendix S3}), we obtain prior parameters $\alpha$=0.998 and $\beta$=4.63. Using these priors in equations~\ref{mean} and~\ref{variance} above (or in the R function ``calculate\_distribution" in \emph{Appendix S3}), we find a prior $\bar\lambda$=0.177 and var($\lambda$)=0.026. Adding the observed data ($n=10$, $k=3$) and using the same code, we obtain posterior parameters $\alpha'$=4.00 and $\beta'$=11.6 and a posterior $\bar\lambda$=0.256 and var($\lambda$)=0.012. Comparing the posterior distribution to the prior, we see that the posterior is closer to the observed data and that the additional data about interactions between species $i$ and $j$ has reduced the variance. We may also wish to calculate a credible interval (analogous to the frequentist confidence interval). This is also quite straightforward in R (see function ``credible\_interval" in \emph{Appendix S3}). In this case, a 95\% credible interval for $\bar\lambda$ is (0.080, 0.491).

        Now, consider the case where the two species have never been observed interacting across $n$ trials, i-e. $k=0$. The question is then ``what is the probability that these two species do not interact''? Since it is not possible to prove that the two species could never interact (strictly speaking, in a Bayesian approach $\lambda=0$ is impossible), we must fix a threshold below which we consider that there is no interaction ($\lambda \sim 0$). We call this threshold probability $\lambda*$. We then use the cumulative distribution function to estimate $P(\lambda<\lambda*|L=0,n)$ for different $n$. The function ``samples\_for\_threshold" in \emph{Appendix S3} calculates distribution function for $\lambda*$ with an increasing number of trials. This yields a surprising result: it requires \textgreater24 observations of no interactions to be 95\% sure that the interaction probability is smaller than $\lambda*$=0.1 (recall Fig.~\ref{upper_limits}, Box 2). Note the special case where there is no observation of the two species co-occurring, $n = 0$.  In this situation, the posterior distribution converges to the prior distribution since the data include no information on the probability with which species might interact should they co-occur.

@@

% Definite second
\section*{Appendix S2: Details of \emph{Salix} data collection}

@@

% Definite third
\section*{Appendix S3: Results for \emph{Salix}-galler component}

@@

      As with the galler-natural enemy component, we estimated frequencies of \emph{Salix}-galler interactions based on the normalised degree of each species (\emph{Appendix S5}). We obtained prior parameters of $\alpha$=8.72, $\beta$=305 for the \emph{Salix}-galler component. These prior parameters were then used to estimate the posterior distribution.

@@

% Definite fourth
\section*{Appendix S4: Analyses using alternative prior}

@@

\section*{Appendix S5: Calculating the credible interval around a probability estimate}

  Here we describe the derivation of the Clopper-Pearson credible interval for the estimated probability of interaction $\lambda$ of a pair of species observed co-occurring $n$ times and interacting $k$ times. As we are most interested in the probability of interaction between species pairs that have never been observed co-occurring, we consider only the case where $k=0$ over a variety of $n$. This is straightforward to do in R (see the function ``credible\_interval" in \emph{Appendix S2}).

@@

\section*{Appendix S7: Calculating posterior distributions and confidence intervals}

@@

>>>>

\section*{Appendix S1: Detailed description of mathematical framework}

  \subsection*{A naive quantification of uncertainty}

    We start by considering how to naively quantify an interaction probability and its associated uncertainty \emph{for an interaction that has not yet been observed}. Consider the case where a pair of species have been observed co-occurring $n$ times, of which they have been observed to interact in $k = 0$ cases. To evaluate the uncertainty of this interaction, consider the occurrence of an interaction as a Bernoulli trial. In this framework, the number of successes $k$ over $n$ trials will follow a binomial distribution:

        \begin{equation}
          L \sim Bin(n,\lambda) ,
        \end{equation}

        \begin{equation}
           P(L = k|\lambda,n) = {n \choose k}\lambda^k(1-\lambda)^{n-k} . 
           \label{likelihood}
        \end{equation}

    \noindent The parameter $\lambda$, the probability of observing an interaction over an infinite time interval and area, is the quantity we want to estimate from empirical data. 
    The maximal likelihood estimate (MLE) of $\lambda$ is straightforward to find given $k$ and $n$:

        \begin{equation}
          \lambda_{MLE} = \frac{k}{n}  .
          \label{theta_MLE}
        \end{equation}

  \subsection*{Posterior distribution of the interaction probability}

    Here we adopt a Bayesian approach to estimate the posterior distribution of the parameter $\lambda$ :

    \begin{equation}
      \underbrace{P(\lambda|k,n)}_{Posterior} = \frac{\overbrace{P(k|\lambda,n)}^{Likelihood}\overbrace{P(\lambda)}^{Prior}}{\underbrace{P(k|N)}_{Normaliser}} .
      \label{posterior}
    \end{equation}

    According to the above description, the likelihood is simply the binomial distribution (Eq.~\ref{likelihood}). Since $\lambda$ is a probability, it is bounded between 0 and 1 and the most appropriate prior distribution is the beta:

    \begin{equation}
      \lambda \sim Beta(\alpha,\beta) , \label{prior}
    \end{equation}

    \noindent which has two shape parameters, $\alpha$ and $\beta$.

   The beta-binomial distribution is a conjugate distribution of the binomial distribution. This allows us to analytically compute the posterior distribution of a binomial model with a beta prior distribution. We can re-write the posterior distribution of $\lambda$ as:

    \begin{equation}
      P(\lambda|k,n) = \frac{\lambda^{\alpha+k-1}(1-\lambda)^{\beta+n-k-1}}{B(\alpha+k,\beta+n-k)} , \label{posterior2}
    \end{equation}

    \noindent where the function $B$ is the beta function. The posterior distribution of $\lambda$ therefore follows the beta distribution with new parameters $\alpha'= \alpha+k$ and $\beta'=\beta+n-k$.

\section*{Appendix S2: Calculating uninformative and informative priors}

@@

        The mean of these interaction probabilities is 0.176, approximately two-thirds the $\lambda_{MLE}$ obtained from the observed data. We can use the distribution of these interaction probabilities as our prior distribution and estimate the uncertainty surrounding our $\lambda_{MLE}$. With some simple R code (function ``calculate\_parameters", \emph{Appendix S7}), we obtain prior parameters $\alpha$=0.998 and $\beta$=4.63. Using these priors in equations~\ref{mean} and~\ref{variance} above (or in the R function ``calculate\_distribution" in \emph{Appendix S7}), we find a prior $\bar\lambda$=0.177 and var($\lambda$)=0.026. Adding the observed data ($n=10$, $k=3$) and using the same code, we obtain posterior parameters $\alpha'$=4.00 and $\beta'$=11.6 and a posterior $\bar\lambda$=0.256 and var($\lambda$)=0.012. Comparing the posterior distribution to the prior, we see that the posterior is closer to the observed data and that the additional data about interactions between species $i$ and $j$ has reduced the variance. We may also wish to calculate a credible interval (analogous to the frequentist confidence interval). This is also quite straightforward in R (see function ``credible\_interval" in \emph{Appendix S7}). In this case, a 95\% credible interval for $\bar\lambda$ is (0.080, 0.491).

        Now, consider the case where the two species have never been observed interacting across $n$ trials, i-e. $k=0$. The question is then ``what is the probability that these two species do not interact''? Since it is not possible to prove that the two species could never interact (strictly speaking, in a Bayesian approach $\lambda=0$ is impossible), we must fix a threshold below which we consider that there is no interaction ($\lambda \sim 0$). We call this threshold probability $\lambda*$. We then use the cumulative distribution function to estimate $P(\lambda<\lambda*|L=0,n)$ for different $n$. The function ``samples\_for\_threshold" in \emph{Appendix S7} calculates distribution function for $\lambda*$ with an increasing number of trials. This yields a surprising result: it requires \textgreater24 observations of no interactions to be 95\% sure that the interaction probability is smaller than $\lambda*$=0.1 (recall Fig.~\ref{upper_limits}, Box 2). Note the special case where there is no observation of the two species co-occurring, $n = 0$.  In this situation, the posterior distribution converges to the prior distribution since the data include no information on the probability with which species might interact should they co-occur.

@@

\section*{Appendix S3: Details of \emph{Salix} data collection}

@@

\section*{Appendix S4: Results for \emph{Salix}-galler component}

@@

      As with the galler-natural enemy component, we estimated frequencies of \emph{Salix}-galler interactions based on the normalised degree of each species (\emph{Appendix S6}). We obtained prior parameters of $\alpha$=8.72, $\beta$=305 for the \emph{Salix}-galler component. These prior parameters were then used to estimate the posterior distribution.

@@

\section*{Appendix S5: Analyses using alternative prior}

@@

@@

\section*{Appendix S7: R code for calculating posterior distributions and confidence intervals}

@@

\section*{Appendix S8: Calculating the credible interval around a probability estimate}

  Here we describe the derivation of the Clopper-Pearson credible interval for the estimated probability of interaction $\lambda$ of a pair of species observed co-occurring $n$ times and interacting $k$ times. As we are most interested in the probability of interaction between species pairs that have never been observed co-occurring, we consider only the case where $k=0$ over a variety of $n$. This is straightforward to do in R (see the function ``credible\_interval" in \emph{Appendix S7}).
